Thesis: draft

### 2021-04-15 draft.tex
latest update

In this chapter the results are shared and discussed. The chapter in subdivided into four main sections
\begin{itemize}
	\item The impact of network structure: the effect of different network topologies on the evolution of the opinions and the formation of echo chambers is reviewed. Main focus lies on networks with a high clustering coefficient and/or a community structure.
	\item The impact of stubborn people: stubborn people are introduced in the system; in this section the effects of those stubborn people on the evolution of opinions and the formation of echo chambers will be discussed.
	\item Random versus community distributed opinions/stubborn agents: this section investigates different ways of distributing the opinions and/or the stubborn agents in a network model with a community structure (SBM).
	\item Real-world networks: the opinion dynamics model of this thesis will be ran on some real-world networks. This section will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
\end{itemize}

@@

	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.

@@

model. This is to be expected, since a stochastic block model with a low modularity, and hence a very blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network. The stochastic Watts-Strogatz block model (SBM-WS) has both a community structure and a high average clustering. The formation of echo chambers when using this model is approximately of the same size compared to the formation of echo chambers in a Watts-Strogatz model with a similar average clustering. This suggests that it is mainly the high clustering that drives the system to form echo chambers, while the community structure plays a less important role. However, as said before, the stochastic block model with a high modularity also has a (smaller) formation of echo chambers, so having a clear and strong community structure definitely plays some role in phenomena such as polarization and the formation of opinion clusters. \\

@@

	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.

@@

This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results (together with the standard deviation $\sigma$) for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.

@@

@@

Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS). For each network, the REC and PR method are compared to each other. The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph.\\

@@

The PR method shows a completely different tendency. There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1. This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1. \\

@@

Networks with a high clustering coefficient (WS and SBM-WS) behave completely different. For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}). An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}. This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity. In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180. These large spikes, indicating a large increase of echo chambers around the minority opinion, don't necessary mean that we have a huge absolute number of echo chambers around the minority opinion 0 at $t=500$. They merely indicate that compared to the beginning at $t=0$ the increase is large. If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm. Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population. Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters. The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm. This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution. In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution. 

@@

Now, lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$. In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted. The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$). This is a measure for the formation of echo chambers (opinion clusters) in the system. The data points represent the value of the echo chamber size in networks with a different modularity. The initial opinion distribution was always $50/50$. \\

@@

\section{Impact of stubborn people}
In this section, the impact of introducing stubborn people will be reviewed. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
\begin{itemize}
	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
\end{itemize}
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.

@@

@@

\textbf{Making a fraction of the nodes completely stubborn}\\
We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$. In all cases, the stubborn nodes are randomly distributed over the network.\\
Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm. The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}. Data points are again shown together with their standard error, $SE$, for the same reasons as explained in Section \ref{echoVSmod}.\\

@@

Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms. Data points are again shown together with their standard error $SE$. The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}). Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease. This is especially clear for the high clustering networks (WS and SBM-WS). The SBM with a high modularity also displays a tiny bump. The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\

@@

\begin{figure}[H]
	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
	\captionsetup{format=plain}
	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb}
\end{figure}

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
    \caption{REF}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \caption{REC}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}
\end{figure}

>>>>

In this chapter the results are shared and discussed. The chapter is subdivided into four main sections
\begin{itemize}
	\item The impact of network structure: the effect of different network topologies on the evolution of the opinions and the formation of echo chambers is reviewed. Main focus lies on networks with a high clustering coefficient and/or a community structure.
	\item The impact of stubborn people: stubborn people are introduced in the system; in this section the effects of those stubborn people on the evolution of opinions and the formation of echo chambers will be discussed.
	\item Random versus community distributed opinions/stubborn agents: this section investigates different ways of distributing the opinions and/or the stubborn agents in a network model with a community structure (SBM).
	\item Real-world networks: the opinion dynamics model of this thesis will be ran on some real-world networks. This section will discuss the results and compare them to results obtained from network models with a similar average degree, clustering coefficient and/or community structure.
\end{itemize}

@@

	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.

@@

model. This is to be expected, since a stochastic block model with a low modularity, and hence a very blurred, almost non-existing community structure, strongly resembles an Erd\H{o}s-R\'{e}nyi network. The stochastic Watts-Strogatz block model (SBM-WS) has both a community structure and a high average clustering. The formation of echo chambers when using this model is approximately of the same size as the formation of echo chambers in a Watts-Strogatz model with a similar average clustering. This suggests that it is mainly the high clustering that drives the system to form echo chambers, while the community structure plays a less important role. However, as said before, the stochastic block model with a high modularity also has a (smaller) formation of echo chambers, so having a clear and strong community structure definitely plays some role in phenomena such as polarization and the formation of opinion clusters. \\

@@

	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.

@@

This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. The results that they obtained are, however, for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results (together with the standard deviation $\sigma$) for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.030$ in Figure \ref{ws10000}.

@@

    \label{er_echo_20-80}

@@

Figure \ref{echo_20_80} shows the distribution of the fraction of friends of node $i$ with the same opinion 1 at $t=500$ for the four network models: Erd\H{o}s-R\'{e}nyi model (ER), Watts-Strogatz model (WS), stochastic block model (SBM) and stochastic Watts-Strogatz block model (SBM-WS). For each network, the REC and PR method are compared to each other. The interpretation of the figure is the same as was described in Subsection \ref{50-50} in the "Echo chambers" paragraph. Error bars are again omitted for a better visualization.\\

@@

The PR method shows a completely different tendency. There is an increase in the fraction of nodes with (almost) all their neighbors carrying the majority opinion 1. This is a consequence from the fact that the PR method reduces the visibility of the minority opinion 0 in the neighborhood of nodes with the majority opinion 1. This reduced visibility of the minority opinion 0 in these neighborhoods drives them eventually to adopt opinion 1, with an increase in the number of echo chambers around the majority opinion 1 as a consequence. Remember from Figure \ref{ev_op_20_80}, that the PR filtering algorithm strongly reduces the prevalence of the minority opinion. Hence, the increase in the formation of echo chambers around the majority opinion 1 is not only a consequence of a reorganization of the system from a random opinion distribution to a clustered opinion distribution, but also from a shift in the initial $20/80$ opinion distribution to a final (close to) $0/100$ opinion distribution. This can also be seen in the fact that, across the $x$-axis, the average number of neighborhoods with some fraction of the nodes carrying the minority opinion 0 becomes equal to zero. \\

@@

Networks with a high clustering coefficient (WS and SBM-WS) behave completely different. For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}). An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}. This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity. In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180. These large spikes, indicating a large increase of echo chambers around the minority opinion, don't necessary mean that we have a huge absolute number of echo chambers around the minority opinion 0 at $t=500$. They merely indicate that, compared to the beginning at $t=0$, the increase is large (and if there are zero echo chambers around the minority opinion 0 at $t=0$, the increase would be infinite even if there is only one echo chamber around the minority opinion 0 at $t=500$). If we would look at the absolute values of the number of echo chambers at $t=500$, we would still find that the vast majority of the like-minded neighborhoods carry the majority opinion 1 and that only a minor fraction carries the minority opinion 0.\\
This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm. Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population for the high clustering networks. Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters. The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm. This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution. In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution. 

@@

Lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$. In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted. The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$). This is a measure for the formation of echo chambers (opinion clusters) in the system. The data points represent the value of the echo chamber size in networks with a different modularity. The initial opinion distribution was always $50/50$. \\

@@

@@

\section{Impact of stubborn people}
In this section, the impact of introducing stubborn people will be reviewed. We focus on two different stubbornness models, both are thoroughly discussed in Chapter \ref{chap3}, Section \ref{stubbBin}
	\item Including a stubbornness parameter: these results are discussed in Section \ref{stubbpar},
	\item The majority threshold model: results for this model are given in Section \ref{majThres}.
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$)
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$; the average clustering coefficient takes values around $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r = 0$ to completely stubborn, $r = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
Once again, the opinion dynamics model described in Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented with an activation probability equal to $p_{\text{act}} = 0.1$ and a maximum length of the actual time-line $S = 20$.

\newpage
\noindent
\textbf{Making a fraction of the nodes completely stubborn}\\
\newline
We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$. In all cases, the stubborn nodes are randomly distributed over the network.\\
	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_stubbornness_SBM_frac_compl_stubb_PR_12x10.png}
	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb}

@@

Figure \ref{echo_vs_frac_complStubb} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the PR filtering algorithm and for an initial $50/50$ opinion distribution. The $y$-axis represents the formation of echo chambers in the same way as was explained for Figure \ref{echo_vs_mod} in Section \ref{echoVSmod}. Data points are again shown together with their standard error, $SE$, for the same reasons as explained in Section \ref{echoVSmod}.\\

@@

Figure \ref{echo_vs_frac_complStubb_REF-REC} shows the formation of echo chambers in the system versus the fraction of completely stubborn nodes, fracRes, when using the REF and REC filtering algorithms. The results are again obtained for an initial $50/50$ opinion distribution. Data points are shown together with their standard error $SE$. The results for the REF and REC method behave qualitatively the same, but are completely different from those obtained with the PR filtering algorithm (Figure \ref{echo_vs_frac_complStubb}). Instead of a decrease in the formation of echo chambers with an increasing fraction of completely stubborn nodes, we can first observe an increase in the formation of echo chambers, followed by a decrease. This is especially clear for the high clustering networks (WS and SBM-WS). The SBM with a high modularity also displays a tiny bump. The ER and the SBM with a low modularity (data points for this model lie, once again, behind the ER data points) do not seem to have a significant bump.\\

@@

  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REF_8x7.png}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_stubbornness_frac_compl_stubb_REC_8x7.png}
    \label{REC_frac_compl_stubb}
  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_frac_complStubb_REF-REC}
Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn nodes ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion 0 is the minority opinion, opinion 1 is the majority opinion). Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm. The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure. The interpretation of the $y$-axis is slightly different than before. In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. Thus we first normalized the echo chambers around opinion 0 and those around opinion 1 and than took the average of both. In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity. We still want to have some average of the echo chambers around opinion 0 and those around opinion 1 to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities. Hence, we chose for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before). Note that by taking the average, we neglect the skewed echo chamber distribution. This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn people and are less interested in the distribution of those echo chambers. \\
When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn nodes. This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm as depicted in Figure \ref{echo_vs_frac_complStubb}. Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn nodes equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}. This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}. The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher. It is however not recommended to compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion 0 and those around opinion 1 in case of an initial $20/80$ opinion distribution. The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}. When comparing the different network models, these skewed distributions should not be overlooked. The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn people and not to compare the different models against each other. \\
% explain reason behavior REC!!
The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, look again different than those for the PR filtering algorithm. The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn nodes, followed by a decrease. This result is very similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}. The stochastic block model with a high modularity doesn't really show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn nodes smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn agents increases further, we can start to see a decrease in the formation of echo chambers. The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn agents (echo chamber size around $1.2$). This very limited, almost non-significant formation of echo chambers for the ER model and the SBM with a low modularity, in combination with the REC filtering algorithm, is also visible in Figure \ref{er_echo_20-80} and Figure \ref{sbm}. For an increasing fraction of the completely stubborn people, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\
We can conclude that the formation of echo chambers versus the fraction of completely stubborn people, for an initial $20/80$ opinion distribution, qualitatively resembles the results for an initial $50/50$ opinion distribution (decrease in formation of echo chambers with an increasing fraction of stubborn agents for the PR method versus an increase in the formation of echo chambers with an increasing fraction of stubborn people, followed by a decrease for the REC method).
